%\thispagestyle{empty}

This thesis is the result of two years of work under the supervision of prof. Sabrina Sartori at the University of Oslo, in the period August 2018 to March 2020. 

First, I would like to thank my supervisors Dr. Sabrina Sartori, Tor Svendsen Bjørheim, Dr. George E. Froudakis and Dr. George Fanourgakis. Sabrina, thank you for all the help throughout my masters thesis, your inexhaustible patience and resourcefulness when we discovered a new problem. Tor, thank you for insight  regarding DFT and thoughts on direction. George Froudakis, thank you for welcoming me to Greece with open arms and a open heart. Thanks to you I will always feel at home on the island of Crete. Lastly, thank you George Fanourgakis, you have been one of the greatest contributor for my success, and thank you for sharing your wast ocean of knowledge on chemoinformatics. 

The University of Oslo and the University of Crete are filled with people that have helped me. I thank you all. Especially, Professor Morten Hjorth-Jensen for igniting my interest for computational physics during his courses at UiO, and later when discussing my thesis. Also, thank you Manolis Tylianakis for always welcoming me with a smile at UoC and being the first to propose a solution.

I would like to thank all the friends that I have made over the past years while studying at the university. Especially, I would like to thank Blindern Studenterhjem for being a great place for students, and all the good friends I have made there. 

Lastly, I would like to thank the people that gave me good laughs and comfort in this period, motivated me to work harder,  and helped me, both with spell checking and fixing the errors that was relevant for the work. Thank you Embla Maria O'cadiz Gustad, Mikael Kiste, Mari Røsvik and Erik Lund. Thank you Ida Grøn Roepstorff and Erik Skaar, your help has been immensely appreciated. 




